---
title: "01-abramovitz"
format: gfm
execute: 
  warning: false
  message: false
---

Here's a quick exploration of some abramovitz data. This is ~pseudo-close~ to Abramovitz' actual model. Here, the incumbent party's voteshare (not two-party voteshare!!!) is estimated by the incumbent president's approval, whether or not the incumbent is running, and the presence of a third party in the race. 

$$
\begin{align*}
\text{Inc. Voteshare} & = \alpha + \beta_1 \ \text{Inc. Net Appr.} + \beta_2 \ \text{Inc. Running} + \beta_3 \ \text{Third Party}
\end{align*}
$$

```{r}
library(tidyverse)
library(riekelib)

abramovitz <- read_csv("../../data/abramovitz.csv")

model_data <- 
  abramovitz %>%
  select(year, 
         inc_party,
         inc_running,
         dem,
         rep,
         inc_approval = fte_net_inc_approval,
         third_party = third_party_present) %>%
  mutate(inc_vote = if_else(inc_party == "dem", dem, rep),
         inc_approval = inc_approval/100)

glm(inc_vote ~ inc_approval + inc_running + third_party,
    family = quasibinomial(),
    data = model_data) %>%
  broom::augment(type.predict = "response",
                 se_fit = TRUE,
                 newdata = crossing(inc_approval = seq(from = -1, to = 1, by = 0.1),
                                    inc_running = 0:1,
                                    third_party = 0:1)) %>%
  mutate(lower = .fitted - .se.fit,
         upper = .fitted + .se.fit) %>%
  ggplot(aes(x = inc_approval,
             y = .fitted)) + 
  geom_ribbon(aes(ymin = lower,
                  ymax = upper,
                  fill = as.character(inc_running)),
              alpha = 0.25) + 
  geom_line(aes(color = as.character(inc_running))) + 
  facet_wrap(~third_party)
```

This is a bit more involved. Using Democrats as a test case, this model looks at the party voteshare based on an interaction between the incumbent party and approval and whether or not a third party is present.

$$
\text{Dem Voteshare} = \alpha + \beta_1 \ \text{Inc. Party} + \beta_2 \ \text{Inc. Net Appr.} + \beta_3 \ \text{Inc. Party} \times \text{Inc. Net Appr.} + \beta_4 \ \text{Third Party}
$$

```{r}
model_data <- 
  abramovitz %>%
  select(dem,
         rep,
         inc_party, 
         inc_running,
         inc_approval = fte_net_inc_approval,
         third_party = third_party_present) %>%
  mutate(inc_approval = inc_approval/100)

glm(dem ~ inc_party*(inc_approval) + third_party,
    family = quasibinomial(),
    data = model_data) %>% 
  broom::augment(newdata = crossing(inc_party = c("dem", "rep"),
                                    inc_approval = seq(from = min(model_data$inc_approval), 
                                                       to = max(model_data$inc_approval), 
                                                       by = 0.01),
                                    third_party = 0:1),
                 type.predict = "response",
                 se_fit = TRUE) %>%
  mutate(lower = .fitted - .se.fit,
         upper = .fitted + .se.fit) %>%
  ggplot(aes(x = inc_approval,
             y = .fitted)) +
  geom_ribbon(aes(ymin = lower,
                  ymax = upper,
                  fill = inc_party),
              alpha = 0.25) + 
  geom_line(aes(color = inc_party)) +
  geom_point(data = model_data,
             mapping = aes(x = inc_approval,
                           y = dem,
                           color = inc_party)) + 
  scale_color_brewer(palette = "Set1", direction = -1) +
  scale_fill_brewer(palette = "Set1", direction = -1) + 
  riekelib::scale_xy_percent() + 
  facet_wrap(~paste0("third_party: ", third_party)) +
  riekelib::theme_rieke() +
  labs(title = glue::glue("**Democratic Voteshare in the Presence of Third Parties**"),
       x = "Incumbent Net Approval",
       y = "Democratic Voteshare")
```

Alrighty, here's Abramovitz' actual model (w/o adjustment for partisanship):

$$
\text{Inc. 2PV} = \alpha + \beta_1 \ \text{Inc. Net App.} + \beta_2 \ \text{Q2GDP Growth} + \beta_3 \ \text{Inc. Running}
$$

```{r}
# blegh - figure out a better way to read this in...
gdp <- read_csv("https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=1317&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=GDP&scale=left&cosd=1947-01-01&coed=2023-01-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Quarterly&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2023-05-03&revision_date=2023-05-03&nd=1947-01-01")

cpi <- read_csv("https://fred.stlouisfed.org/graph/fredgraph.csv?bgcolor=%23e1e9f0&chart_type=line&drp=0&fo=open%20sans&graph_bgcolor=%23ffffff&height=450&mode=fred&recession_bars=on&txtcolor=%23444444&ts=12&tts=12&width=1317&nt=0&thu=0&trc=0&show_legend=yes&show_axis_titles=yes&show_tooltip=yes&id=CPIAUCSL&scale=left&cosd=1947-01-01&coed=2023-03-01&line_color=%234572a7&link_values=false&line_style=solid&mark_type=none&mw=3&lw=2&ost=-99999&oet=99999&mma=0&fml=a&fq=Monthly&fam=avg&fgst=lin&fgsnd=2020-02-01&line_index=1&transformation=lin&vintage_date=2023-05-03&revision_date=2023-05-03&nd=1947-01-01")

max_cpi <- max(cpi$CPIAUCSL)

gdp %>%
  left_join(cpi) %>%
  rename_with(str_to_lower) %>%
  rename(cpi = cpiaucsl) %>%
  mutate(gdp_adjust = gdp*max_cpi/cpi) %>%
  pivot_longer(starts_with("gdp")) %>%
  ggplot(aes(x = date, 
             y = value,
             color = name)) + 
  geom_line() +
  scale_color_manual(values = c("gray60", "black")) + 
  theme_rieke() +
  scale_y_comma() +
  labs(x = NULL,
       y = NULL)
```

```{r}
gdp %>%
  left_join(cpi) %>%
  rename_with(str_to_lower) %>%
  rename(cpi = cpiaucsl) %>%
  mutate(gdp_adjust = gdp*max_cpi/cpi,
         gdp_growth = (gdp_adjust - lag(gdp_adjust, n = 4))/lag(gdp_adjust, n = 4)) %>%
  ggplot(aes(x = date,
             y = gdp_growth)) + 
  geom_line() +
  scale_y_percent() +
  theme_rieke()
```

```{r}
model_gdp <- 
  gdp %>%
  left_join(cpi) %>%
  rename_with(str_to_lower) %>%
  rename(cpi = cpiaucsl) %>%
  mutate(gdp_adjust = gdp*max_cpi/cpi,
         gdp_growth = (gdp_adjust - lag(gdp_adjust, n = 4))/lag(gdp_adjust, n = 4),
         quarter = quarter(date),
         year = year(date)) %>%
  filter(quarter == 3) %>%
  select(year, gdp_growth)

model_data <- 
  abramovitz %>%
  left_join(model_gdp) %>%
  mutate(inc_2pv = if_else(inc_party == "dem", dem/(dem + rep), rep/(dem + rep))) %>%
  select(inc_2pv,
         inc_running,
         inc_approval = fte_net_inc_approval,
         gdp_growth,
         year) %>%
  mutate(inc_approval = inc_approval/100)

model_data %>%
  pivot_longer(c(inc_approval, gdp_growth),
               names_to = "x_var",
               values_to = "x") %>%
  ggplot(aes(x = x,
             y = inc_2pv)) +
  geom_point() +
  facet_wrap(~x_var, scales = "free_x") +
  theme_rieke() + 
  scale_xy_percent()
```

```{r}
glm(inc_2pv ~ inc_approval + gdp_growth + inc_running,
    family = quasibinomial(),
    data = model_data) %>%
  broom::augment(type.predict = "response") %>%
  left_join(model_data) %>%
  ggplot(aes(x = inc_2pv, 
             y = .fitted,
             label = year)) + 
  geom_text() + 
  geom_abline(linetype = "dashed") +
  theme_rieke() + 
  scale_xy_percent()
```

```{r}
glm(inc_2pv ~ inc_approval + gdp_growth + inc_running,
    family = quasibinomial(),
    data = model_data) %>%
  broom::augment(type.predict = "response",
                 se_fit = TRUE,
                 newdata = crossing(inc_approval = seq(from = min(model_data$inc_approval),
                                                       to = max(model_data$inc_approval),
                                                       length.out = 30),
                                    gdp_growth = seq(from = -0.05,
                                                     to = 0.05,
                                                     length.out = 3),
                                    inc_running = 0:1)) %>%
  mutate(inc_running = paste0("incumbent running: ", inc_running),
         gdp_growth = paste0("gdp growth: ", scales::label_percent()(gdp_growth)),
         lower = .fitted - .se.fit,
         upper = .fitted + .se.fit) %>%
  ggplot(aes(x = inc_approval,
             y = .fitted)) +
  geom_ribbon(aes(ymin = lower,
                  ymax = upper,
                  fill = inc_running),
              alpha = 0.25) + 
  geom_line(aes(color = inc_running)) +
  geom_hline(yintercept = 0.5,
             linetype = "dashed",
             alpha = 0.5) +
  geom_vline(xintercept = 0,
             linetype = "dashed",
             alpha = 0.5) + 
  facet_wrap(~gdp_growth) +
  scale_xy_percent() + 
  theme_rieke() +
  theme(legend.position = "top",
        legend.title = element_blank())
```

