---
title: "Output Review"
subtitle: "Diagnositc summaries and weekly changelogs"
format: gfm
execute: 
  echo: false
  message: false
  warning: false
---

```{r}
library(tidyverse)
library(riekelib)
```

This document contains a host of information/metadata about the model runtime and sampler conditions that is useful but not appropriate for displaying on-site. There are three broad categories of information displayed:

**Weekly updates**

* Top-level change in the probability of winning in the past week
* State-level ratings changes

**Data**

* Number of polls conducted by rundate
* Number of state-level polls by probability of a democratic win
* Biden's forecasted approval by rundate

**Diagnostics**

* Observations by rundate
* Model runtime by rundate
* Number of divergent transitions by rundate
* Number of transitions at max treedepth by rundate

I also include here a plot of the change in state ratings over time because I think it looks nifty.

## Weekly updates

Weekly change in probability of winning the presidency:

```{r}
#| results: asis
toplines <- 
  read_csv("polls/win_pres.csv") %>% 
  left_join(read_csv("polls/tie_pres.csv")) %>%
  filter(run_date %in% c(Sys.Date(), Sys.Date() - 7)) %>%
  mutate(p_lose = 1 - p_win - p_tie) %>%
  arrange(run_date)

glue::glue(
  "* Harris: **{scales::label_percent(accuracy = 1)(toplines$p_win[1])}** &rarr; **{scales::label_percent(accuracy = 1)(toplines$p_win[2])}**",
  "* Trump: **{scales::label_percent(accuracy = 1)(toplines$p_lose[1])}** &rarr; **{scales::label_percent(accuracy = 1)(toplines$p_lose[2])}**",
  .sep = "\n"
)
```

Weekly change in state ratings:

```{r}
read_csv("polls/win_state.csv") %>%
  filter(state != "National",
         run_date %in% c(Sys.Date(), Sys.Date() - 7)) %>%
  mutate(rating = case_when(p_win > 0.99 | p_win < 0.01 ~ "Safe",
                            p_win > 0.85 | p_win < 0.15 ~ "Very likely",
                            p_win > 0.65 | p_win < 0.35 ~ "Likely",
                            .default = "Uncertain"),
         leader = if_else(p_win > 0.5, "Dem", "Rep"),
         rating = if_else(rating != "Uncertain", paste(rating, leader), rating),
         run_date = if_else(run_date == Sys.Date(), "today", "last_week")) %>%
  select(state, run_date, rating) %>%
  pivot_wider(names_from = run_date,
              values_from = rating) %>%
  filter(today != last_week) %>%
  knitr::kable()
```

```{r}
pal <-
  c(
    "#3579ac",
    "#7cb0d7",
    "#d3e5f2",
    "#f2f2f2",
    "#f2d5d5",
    "#d78080",
    "#b13737"
  )

read_csv("polls/win_state.csv") %>%
  left_join(read_csv("../data/static/electors.csv")) %>%
  mutate(rating = case_when(p_win < 0.01 ~ "Safe Rep",
                            p_win < 0.15 ~ "Very likely Rep",
                            p_win < 0.35 ~ "Likely Rep",
                            p_win < 0.65 ~ "Uncertain",
                            p_win < 0.85 ~ "Likely Dem",
                            p_win < 0.99 ~ "Very likely Dem",
                            TRUE ~ "Safe Dem")) %>%
  filter(state != "National") %>%
  group_by(run_date, rating) %>%
  summarise(electors = sum(electors)) %>%
  ungroup() %>%
  mutate(rating = fct_relevel(rating, c("Safe Rep", 
                                        "Very likely Rep", 
                                        "Likely Rep",
                                        "Uncertain", 
                                        "Likely Dem", 
                                        "Very likely Dem", 
                                        "Safe Dem"))) %>%
  ggplot(aes(x = run_date,
             y = electors,
             fill = rating)) + 
  geom_col(width = 1) +
  scale_fill_manual(values = pal[7:1]) +
  theme_void() +
  theme(legend.position = "none") + 
  expand_limits(x = mdy(c("5/1/24", "11/5/24")))
```


## Data

```{r}
read_csv("polls/polls_out.csv") %>%
  filter(dem_candidate == "Harris") %>%
  count(end_date) %>%
  arrange(end_date) %>%
  mutate(n = cumsum(n)) %>%
  ggplot(aes(x = end_date,
             y = n)) + 
  geom_step() +
  scale_y_comma() + 
  theme_rieke() + 
  labs(title = "**Pollapalooza**",
       subtitle = "Number of Harris/Trump polls conducted by poll end date",
       x = NULL,
       y = NULL) +
  expand_limits(x = mdy("5/1/24"))
```

```{r}
read_csv("polls/polls_out.csv") %>%
  filter(state != "National",
         dem_candidate == "Harris") %>%
  count(state) %>%
  left_join(read_csv("polls/win_state.csv") %>% filter(run_date == max(run_date))) %>%
  left_join(read_csv("../data/static/state_abbreviations.csv")) %>%
  ggplot(aes(x = p_win,
             y = n,
             label = abbr)) + 
  ggrepel::geom_text_repel() +
  scale_x_percent() + 
  theme_rieke() +
  labs(title = "**State schenanigans**",
       subtitle = "Number of Harris/Trump polls by p(Harris win)",
       x = NULL,
       y = NULL) +
  expand_limits(x = c(0, 1))
```

```{r}
read_csv("approval/e_day_approval_current.csv") %>%
  select(run_date, mean, sd) %>%
  crossing(p = c(0.66, 0.95)) %>%
  mutate(.lower = qnorm((1 - p)/2, mean, sd),
         .upper = qnorm((1 - p)/2 + p, mean, sd)) %>%
  ggplot(aes(x = run_date,
             y = mean,
             ymin = .lower,
             ymax = .upper,
             group = p)) + 
  geom_ribbon(alpha = 0.125) + 
  geom_line() +
  scale_y_percent() +
  theme_rieke() +
  labs(title = "**Forecasting a wide net**",
       subtitle = "Biden's forecasted election-day net approval",
       x = NULL,
       y = NULL,
       caption = paste("Shaded region indicates 66/95% credible interval"))
```

## Diagnostics

```{r}
read_csv("model_log.csv") %>%
  ggplot(aes(x = run_date,
             y = observations)) + 
  geom_point(shape = 21,
             alpha = 0.5) + 
  scale_y_comma() +
  facet_wrap(~model_name, scales = "free_y") +
  theme_rieke() + 
  labs(title = "**The observation deck**",
       subtitle = "Number of observations in each model by run date",
       x = NULL,
       y = NULL) +
  expand_limits(y = c(0))
```

```{r}
read_csv("model_log.csv") %>%
  mutate(run_ts = end_ts - start_ts) %>%
  ggplot(aes(x = run_date,
             y = run_ts)) + 
  geom_point(shape = 21,
             alpha = 0.5) +
  scale_y_comma() + 
  facet_wrap(~model_name, scales = "free_y") +
  theme_rieke() +
  labs(title = "**Runnin' up that hill**",
       subtitle = "Model runtime (s) by run date",
       x = NULL,
       y = NULL) + 
  expand_limits(y = c(0))
```

```{r}
read_csv("model_log.csv") %>%
  ggplot(aes(x = run_date,
             y = num_divergent)) + 
  geom_point(shape = 21,
             alpha = 0.5) +
  scale_y_comma() + 
  facet_wrap(~model_name, scales = "free_y") +
  theme_rieke() +
  labs(title = "**Do or diverge**",
       subtitle = "Number of divergent transitions by run date",
       x = NULL,
       y = NULL) + 
  expand_limits(y = c(0)) 
```

```{r}
read_csv("model_log.csv") %>%
  ggplot(aes(x = run_date,
             y = num_max_treedepth)) + 
  geom_point(shape = 21,
             alpha = 0.5) +
  scale_y_comma() + 
  facet_wrap(~model_name, scales = "free_y") +
  theme_rieke() + 
  labs(title = "**Going for distance**",
       subtitle = "Number of transitions at max treedepth by run date",
       x = NULL,
       y = NULL) + 
  expand_limits(y = c(0))
```

